---
title: 'COMPASS Synoptic CB: AQ600 Water Level Calculation: September 2023'
author: "Stephanie J. Wilson"
date: "2024-05-14"
output: html_document
---
#COMPASS Synoptic CB 
#Aquatroll waterlevel calculations 

Authors: Peter Regier & Stephanie J. Wilson 
Last Updated: 05-07-2024

Goal of this code: 
  To take the raw L1 data and calculate waterlevels from pressure and density measured by the aquatroll overtime 
```{r setup, include=FALSE}

library(googlesheets4)
library(dplyr)
library(ggplot2)
library(tidyr)
library(lubridate)

setwd("S:/Biogeochemistry/People/Wilson (Steph)/Compass N Cycle Experiment/Full Incub. Experiment/TAI_NCycle/COMPASS-FME Supplemental Data")
```

#Read in L1 Dataset
```{r}

#download the L1 dataset from: https://drive.google.com/drive/folders/1RP-hxmjAYNGzTfCNpYbMN6U_zbZmEWHM

#set path to local drive where files can be found: 
L1 <- "S:/Biogeochemistry/People/Wilson (Steph)/Compass N Cycle Experiment/Full Incub. Experiment/Manuscript/Supp Sensor Figs/L1"

#read in datafiles from L1 
files <- list.files(L1, pattern = "*.csv", full.names = TRUE, recursive = TRUE)

```

#Read files and create master dataframe
```{r}
#create dataframe for collated data 
results <- list()

#for(f in files[[150]]){ #example to trial with a single file
for(f in files){
  message(f)
  
  selected_research_names = c("gw_pressure", "gw_density", "gw_temperature", "gw_salinity")
  
  dat <- read.csv(f)
  
  troll <- subset(dat, Instrument == "AquaTROLL600")
  
  gw <- troll %>% 
    filter(research_name %in% selected_research_names) %>%
    dplyr::select(TIMESTAMP, research_name, Value, F_OOB, F_OOS, Plot, Site)
  
  results[[f]] <- gw 
}

#we need to bind these together 
dat <- do.call(rbind, results)
rownames(dat) <- NULL
head(dat)

#convert to wide format 
gw1 <- pivot_wider(dat, id_cols = c("TIMESTAMP", "Plot", "Site"), names_from = research_name, values_from = Value )
head(gw1)

gw1$year <- lubridate::year(gw1$TIMESTAMP)
head(gw1)
```

#Pull in Well Dimension & Depth to Sensor Data
```{r}
## Second, read in well dimensions. These are measurements made when installing
## sensors, and are used to calculate the distance below-ground that the 
well_dimensions <- read_sheet("https://docs.google.com/spreadsheets/d/1O4sHvj2FO7EcWEm3WpKEZhFubGn8HCcUsz9EFXhQTXM/edit#gid=0") %>% 
  mutate(Plot = case_when(transect_location == "Upland" ~ "UP", 
                              transect_location == "Transition" ~ "TR", 
                              transect_location == "Wetland" ~ "W"), 
         ground_to_sensor_cm = ring_to_pressure_sensor_cm - (well_top_to_ground_cm - bolt_to_cap_cm)) %>% 
  dplyr::select(site, Plot, ground_to_sensor_cm)

#Got to change the Column names so that Site is also startign with a captial letter 
#colnames(well_dimensions) <- c("Site", "Plot", "ground_to_sensor_cm")
#head(well_dimensions)

gw1$Plot <- as.factor(gw1$Plot)
levels(gw1$Plot) <- list(UP = "UP", TR = "TR", W = "W", UP = "SWAMP")

## Add well dimensions to dataset: inner_join ensures no uncorrected WL data makes it through
df_raw_depths <- inner_join(gw1, well_dimensions, by = c("Site" = "site", "Plot")) #%>% 
  #filter(!is.na(pressure_mbar))
head(df_raw_depths)

```

#Calculate water depth from pressure, density, and well data
```{r}

#gw_density 
#L1 data bounds: 0.98, 1.05
#units g/cm3

#gw_pressure          
#sensor units: psi      
#L1 data set units: mbar      
#L1 data bounds: -10, 910    
#Info: AQ600 Vented pressure already corrected for barometric pressure

#Convert to water depth (h = P [mbar] * 100 [Pa/mbar])/(rho [g/cm3]*1000 [kg/m3//g/cm3]*g [m/s2]) where [Pa]=[kgm/s2m2]
#9.80665 = specific gravity 

#second step is water level below surface of ground (divide ground to sensor so it is in m) subtract ground to sensor from pressure head

df <- df_raw_depths %>% 
  mutate(pressurehead_m = (gw_pressure * 100) / (gw_density * 1000 * 9.80665), 
         wl_below_surface_m = pressurehead_m - (ground_to_sensor_cm / 100)) 
head(df)


```

#Take an initial look at the data 
```{r}

compass_coolors <- c("#FFBC42", "#20063B",  "#419973")
#should we try to do some initial plots? 
#maybe we subset by week or something and plot some up? 

df$year <- as.factor(df$year)
levels(df$year)
df$day <- lubridate::day(df$TIMESTAMP)
df$datetime = parsedate::parse_date(df$TIMESTAMP)


ggplot(df, aes(datetime, wl_below_surface_m, color = Plot)) +
  geom_point( alpha = 0.5) +
  facet_wrap(year ~ Site, scales = "free_y") + 
  labs(x = "", y = "Waterlevel (m)") +
  scale_x_datetime(date_breaks = "3 days", date_labels = "%m/%d")
  #scale_color_manual(values = location_colors) #+
  #scale_x_datetime(date_breaks = "3 days", date_labels = "%m/%d")
 #facet_wrap(~Site, nrow = 1, scales = "free_y") +

yr22 <- subset(df, year == "2022")
plot22 <- ggplot(yr22, aes(datetime, wl_below_surface_m, color = Plot)) +
           geom_point( alpha = 0.5) +
            geom_line( ) + 
             facet_wrap(~ Site, scales = "free_y") + 
             labs(title = "2022", x = "", y = "Waterlevel (m)") +
            scale_x_datetime(date_breaks = "3 days", date_labels = "%m/%d") + 
             theme_classic() + 
             scale_color_manual(values=compass_coolors) +
              theme(panel.background = element_rect(colour = "black", linewidth  =1.2), 
                  legend.position = "none", 
                   legend.title= element_blank(), 
                axis.text=element_text(size=11), 
                 axis.title=element_text(size=12), 
                strip.text.x = element_text(size = 12)) 


yr23 <- subset(df, year == "2023")
plot23 <- ggplot(yr23, aes(datetime, wl_below_surface_m, color = Plot)) +
           geom_point( alpha = 0.5) +
            geom_line( ) + 
             facet_wrap(~ Site, scales = "free_y") + 
             labs(title = "2023", x = "", y = "Waterlevel (m)") +
            scale_x_datetime(date_breaks = "3 days", date_labels = "%m/%d") + 
             theme_classic() + 
             scale_color_manual(values=compass_coolors) +
              theme(panel.background = element_rect(colour = "black", linewidth  =1.2), 
                  legend.position = "bottom", 
                   legend.title= element_blank(), 
                axis.text=element_text(size=11), 
                 axis.title=element_text(size=12), 
                strip.text.x = element_text(size = 12)) 

```

#Salinity Data 
```{r}

compass_coolors <- c("#FFBC42", "#20063B",  "#419973")
#should we try to do some initial plots? 
#maybe we subset by week or something and plot some up? 

df$year <- as.factor(df$year)
levels(df$year)
df$day <- lubridate::day(df$TIMESTAMP)
df$datetime = parsedate::parse_date(df$TIMESTAMP)


ggplot(df, aes(datetime, gw_salinity, color = Plot)) +
  geom_point( alpha = 0.5) +
  facet_wrap(year ~ Site, scales = "free_y") + 
  labs(x = "", y = "Waterlevel (m)") +
  scale_x_datetime(date_breaks = "3 days", date_labels = "%m/%d")
  #scale_color_manual(values = location_colors) #+
  #scale_x_datetime(date_breaks = "3 days", date_labels = "%m/%d")
 #facet_wrap(~Site, nrow = 1, scales = "free_y") +

sal22 <- ggplot(yr22, aes(datetime, gw_salinity, color = Plot)) +
           geom_point( alpha = 0.5) +
            geom_line( ) + 
             facet_wrap(~ Site, scales = "free_y") + 
             labs(title = "2022", x = "", y = "Salinity") +
            scale_x_datetime(date_breaks = "3 days", date_labels = "%m/%d") + 
             theme_classic() + 
             scale_color_manual(values=compass_coolors) +
              theme(panel.background = element_rect(colour = "black", linewidth  =1.2), 
                  legend.position = "none", 
                   legend.title= element_blank(), 
                axis.text=element_text(size=12), 
                 axis.title=element_text(size=12), 
                strip.text.x = element_text(size = 12)) 


sal23 <- ggplot(yr23, aes(datetime, gw_salinity, color = Plot)) +
           geom_point( alpha = 0.5) +
            geom_line( ) + 
             facet_wrap(~ Site, scales = "free_y") + 
             labs(title = "2023", x = "", y = "Salinity") +
            scale_x_datetime(date_breaks = "3 days", date_labels = "%m/%d") + 
             theme_classic() + 
             scale_color_manual(values=compass_coolors) +
              theme(panel.background = element_rect(colour = "black", linewidth  =1.2), 
                  legend.position = "bottom", 
                   legend.title= element_blank(), 
                axis.text=element_text(size=12), 
                 axis.title=element_text(size=12), 
                strip.text.x = element_text(size = 12)) 


box1 <- ggplot()+
  geom_boxplot(data=yr22, aes(x=Plot, y=gw_salinity, fill=Plot)) + 
  facet_wrap(~Site, ncol = 2) + 
  #facet_grid(~factor(Site, levels=c("SWH", "GCW","MSM", "GWI"))) +
  theme_classic() + 
  labs(title= "2022", x=" ", y= "Salinity") +
  scale_fill_manual(values=compass_coolors) +
  #scale_x_discrete(labels=c("Upland", "Transition", "Wetland")) +
  theme(panel.background = element_rect(colour = "black", linewidth  =1.2), 
        legend.position = "NONE", 
        legend.title= element_blank(), 
        axis.text=element_text(size=12), 
        axis.title=element_text(size=12), 
        strip.text.x = element_text(size = 12)) 

box1
```
